Short and Mechanized Logical Relation for Dependent Type Theory

### 2024-02-24 paper/paper.tex
Answer some questions

With untyped conversion,
we sidestep the relational, Kripke-style logical relation found in
other mechanized proofs. \scw{Need to define Kripke-style. Also the other
proofs need Kripke style because they are defining typed relations, not untyped
relations. }
However, our early dependence on confluence
before the fundamental theorem is established can be alarming.
In a system with type-directed reduction,
confluence is not immediately available because it
depends on $\Pi$-injectivity, which is usually only proven after the
fundamental theorem.\scw{confluence depends on Pi injectivity? I thought
it was only needed for subject reduction}
Fortunately, there are syntactic workarounds for the $\Pi$-injectivity
problem that allow us to recover confluence property independently
from the logical relation. \citet{siles2012pure} generalize the
notion of Type Parallel One Step Reduction from \citet{adams2006pure}
to syntactically prove $\Pi$-injectivity for arbitrary Pure Type
Systems. \citet{weirich:systemd} add $\Pi$-injectivity to their
equational theory, thus allowing subject reduction to be proven
independently from confluence. Therefore, we believe that even in a
system where type-directed reduction is required (e.g. a system with
the unit $\eta$-law) in the logical relation, the proof can still be carried
out in a structure similar to the one we have presented. \scw{Not sure that
I understand this paragraph}

>>>>

With untyped conversion,
we sidestep the relational, Kripke-style logical relation found in
other mechanized proofs. \scw{Need to define Kripke-style. Also the other
proofs need Kripke style because they are defining typed relations, not untyped
relations. }
However, our early dependence on confluence
before the fundamental theorem is established can be alarming.
In a system with type-directed reduction,
confluence is not immediately available because it
depends on $\Pi$-injectivity, which is usually only proven after the
fundamental theorem.\scw{confluence depends on Pi injectivity? I thought
it was only needed for subject reduction}\yl{it's
transitive. Confluence depends on subject reduction, which in turn
depends on pi injectivity. Maybe it's worth spelling out the details}
Fortunately, there are syntactic workarounds for the $\Pi$-injectivity
problem that allow us to recover confluence property independently
from the logical relation. \citet{siles2012pure} generalize the
notion of Type Parallel One Step Reduction from \citet{adams2006pure}
to syntactically prove $\Pi$-injectivity for arbitrary Pure Type
Systems. \citet{weirich:systemd} add $\Pi$-injectivity to their
equational theory, thus allowing subject reduction to be proven
independently from confluence. By adopting these techniques that allow
us to derive confluence early even for systems with type-directed
reduction, we believe our proof technique can significantly shorten
the existing logical relation proofs for systems with typed judgmental
equality and more complex $\eta$ laws. We leave that as part of our
future work.
% Therefore, we believe that even in a
% system where type-directed reduction is required (e.g. a system with
% the unit $\eta$-law) in the logical relation, the proof can still be carried
% out in a structure similar to the one we have presented.
\scw{Not sure that I understand this paragraph}
\yl{Reworded slightly to emphasize it's future work that we haven't
  done and it is speculative}
